---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
library(dplyr)
library(tidyr)
library(readr)

# Select only relevant columns
deaths <- av %>%
  select(URL, starts_with("Death")) %>%
  pivot_longer(
    cols = starts_with("Death"),
    names_to = "Time",
    values_to = "Death"
  ) %>%
  mutate(
    Time = parse_number(Time),   
    Death = tolower(Death)       
  )

head(deaths)

returns <- av %>%
  select(URL, starts_with("Return")) %>%
  pivot_longer(
    cols = starts_with("Return"),
    names_to = "Time",
    values_to = "Return"
  ) %>%
  mutate(
    Time = parse_number(Time),
    Return = tolower(Return)
  )

head(returns)

avg_deaths <- deaths %>%
  group_by(URL) %>%
  summarize(num_deaths = sum(Death == "yes", na.rm = TRUE)) %>%
  summarize(avg_deaths = mean(num_deaths))

avg_deaths
```

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers. 

## Individually
## Kenzie's Work
For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.
"If all the Avengers are sent to a pocket dimension after seemingly being killed by Onslaught but we then we follow their adventures on Counter-Earth, they’re not dead."
### Include the code

Make sure to include the code to derive the (numeric) fact for the statement
```{r}
# Load required packages
library(dplyr)
library(tidyr)
library(readr)

# Load the Avengers dataset
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
# Create deaths dataset in long format
deaths <- av %>%
  select(URL, Name.Alias, starts_with("Death")) %>%
  pivot_longer(
    cols = starts_with("Death"),
    names_to = "Time",
    values_to = "Death"
  ) %>%
  mutate(
    Time = parse_number(Time),
    Death = tolower(Death)
  )

# Create returns dataset in long format
returns <- av %>%
  select(URL, Name.Alias, starts_with("Return")) %>%
  pivot_longer(
    cols = starts_with("Return"),
    names_to = "Time",
    values_to = "Return"
  ) %>%
  mutate(
    Time = parse_number(Time),
    Return = tolower(Return)
  )

# Merge deaths and returns
death_return <- left_join(deaths, returns, by = c("URL", "Name.Alias", "Time"))

# Identify Avengers who died and then returned
revived <- death_return %>%
  filter(Death == "yes", Return == "yes") %>%
  distinct(Name.Alias)

# Onslaught-era Avengers involved in the pocket dimension storyline
onslaught_avengers <- c("Captain America", "Iron Man", "Thor", "Scarlet Witch")

# Filter to see which of them both died and returned
revived_onslaught <- revived %>%
  filter(Name.Alias %in% onslaught_avengers)

# Derive a numeric fact: How many Onslaught-era Avengers returned?
num_revived <- nrow(revived_onslaught)
num_revived
```
### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.
"Based on the dataset, three of the four heroes involved in the storyline are reported as dying and then returning. This shows that they are not dead, but continuing their adventures on the counter-earth. Those three Avengers are Thor, Captain America, and Iron Man.
Upload your changes to the repository. Discuss and refine answers as a team.
